SST_patterns

### 2019-09-20 SST_patterns.Rmd
updated
The pixel filtering by distance now functions as expected

library(tidyverse)
library(lubridate)
library(ggpubr)
library(zoo)
library(FNN)
library(scales)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
# library(ncdf4) # This was used to process NetCDF files in an earlier version
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ncdf4) # library for processing netCDFs
library(ncdf4)
library(data.table)
library(plyr)
```

@@

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData") # 5 decimal places
load("Data/OISST.RData") # 2 decimal places
OISST <- BC_avhrr_only_v2_Document_Document 
rm(BC_avhrr_only_v2_Document_Document )
load("Data/CMC.RData") # 1decimal places
```

@@

```{r}
# RWS: It appears as though the satellite data you are loading here have already been trimmed down to the one nearest pixel to each SACTN site
# This won't work for the following steps because you need to be finding pixels from the COMPLETE set of satellite data you downloaded
# Once you load the complete satellite data lon/lat values you can run the following code and it should give you the coordinates you want

site_pixels$lat<-round(site_pixels$lat, 1) #CMC is to 1 decimal place and OISST is to 2
site_pixels$lon<-round(site_pixels$lon, 1)

# These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
# MUR <- MUR %>% 
#   select(lon, lat) %>% 
#   mutate(product = "MUR")

# Decided to work with OISST and CMC as both has a time series of 30years

OISST_prod <- OISST %>% 
  select(lon, lat) %>% 
  mutate(product = "OISST")

CMC_prod <- CMC %>% 
  select(lon, lat) %>% 
  mutate(product = "CMC")

sat_data <- rbind(CMC_prod, OISST_prod) %>% 
  #rbind(., CMC) %>% 
  select(product, lon, lat)

# rm(MUR, OISST, CMC); gc()

sat_pixels <- sat_data %>% 
  select(product, lon, lat) %>%
  unique()
## For testing the nest/map pipeline
# df <- sat_pixels %>% 
#   filter(product == "MUR", site == "Lamberts Bay") %>% 
#   select(-c(product, site))

match_func <- function(df){
  df <- na.omit(df) %>% 
    dplyr::rename(lon_sat = lon, lat_sat = lat)
  match_index <- knnx.index(data = as.matrix(round(site_pixels[,2:3]),1),
                    query = as.matrix(df[,1:2]), k = 1)
  res <- cbind(site_pixels[match_index,], df)
  return(res)
}

pixel_match <- sat_data %>% 
  group_by(product) %>% 
  nest() %>% 
  mutate(match = purrr:::map(data, match_func)) %>% 
  select(-data) %>%
  unnest(match) #%>% 
  #unite(col = combi, lon, lat, sep = "/", remove = F)

pixel_match$lat<-round(pixel_match$lat, 1) #CMC is to 1 decimal place and OISST is to 2
pixel_match$lon<-round(pixel_match$lon, 1)

# You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
# one time series for each product for each pixel at each distance for each site that you may then
# dataset_filt <- function(df){
#   filtered_products <- df %>% 
#   unite(col = combi, lon, lat, sep = "/", remove = F) #%>% 
#   #filter(combi %in% pixel_match$combi)
# }

# OISST_fil <- dataset_filt(df = OISST)
# CMC_fil <- dataset_filt(df = CMC)
# 
# 
# pixel_match$lat
# pixel_match$lon

filterted_product <- function(df){
  extracted <- pixel_match %>% 
  right_join(df, by = c("lon", "lat")) %>% 
  select(date,temp,lat,lon)
}

OISST_fil <- filterted_product(df = OISST)
CMC_fil <- filterted_product(df = CMC)

# Next to run upwelling index
```

>>>>

library(tidyverse)
library(lubridate)
library(ggpubr)
library(zoo)
library(FNN)
library(scales)
library(gridExtra)
library(circular)
library(fossil)
library(mapproj)
# library(ncdf4) # This was used to process NetCDF files in an earlier version
library(stringr)
library(doMC); doMC::registerDoMC(cores = 4)
library(fasttime)
library(xtable)
library(ncdf4) # library for processing netCDFs
library(data.table)
# library(plyr) # RWS: Never load the plyr package as it interferes with the tidyverse
library(heatwaveR)
```

@@

```{r}
load("Data/site_list_sub.Rdata")
load("Data/SACTN_US.RData")
load("Data/site_pixels.RData") # 5 decimal places
load("Data/OISST.RData") # 2 decimal places
OISST <- BC_avhrr_only_v2_Document_Document 
rm(BC_avhrr_only_v2_Document_Document ); gc()
load("Data/CMC.RData") # 1decimal places
```

@@

```{r}
# These following three objects (MUR, OISST, CMC) need to be the complete set of lon/lat values for your satellite data
# MUR <- MUR %>% 
#   select(lon, lat) %>% 
#   mutate(product = "MUR")

# Decided to work with OISST and CMC as both has a time series of 30years
OISST_prod <- OISST %>% 
  select(lon, lat) %>% 
  unique() %>% 
  mutate(product = "OISST")

CMC_prod <- CMC %>% 
  select(lon, lat) %>%
  unique() %>% 
  mutate(product = "CMC")

# sat_data <- rbind(CMC_prod, OISST_prod) %>% 
  #rbind(., CMC) %>% 
  # select(product, lon, lat)

# sat_pixels <- sat_data %>% 
  # select(product, lon, lat) %>%
  # unique()

## For testing the nest/map pipeline
# df <- site_pixels %>%
#   filter(site == "Lamberts Bay") %>%
#   select(-site)

match_func <- function(df){
  df <- df %>%
    dplyr::rename(lon_site = lon, lat_site = lat)
  OISST_index <- OISST_prod[as.vector(knnx.index(as.matrix(OISST_prod[,c("lon", "lat")]),
                                                 as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
    cbind(., df)
  CMC_index <- CMC_prod[as.vector(knnx.index(as.matrix(CMC_prod[,c("lon", "lat")]),
                                             as.matrix(df[,c("lon_site", "lat_site")]), k = 1)),] %>% 
    cbind(., df)
  res <- rbind(OISST_index, CMC_index)
  return(res)
}

# Find the nearest pixels to each spot along the transect
# NB: SOme pixels are used more thanonce in a transect as the spacing isn't quite larger enough between transect
# points to not fall inside of the same 25 KM pixel
pixel_match <- site_pixels %>% 
  group_by(site) %>% 
  group_modify(~match_func(.x))

# You may then use the 'pixel_match' object to filter out the desired pixels from the full satellite products
OISST_fill <- right_join(OISST, filter(pixel_match, product == "OISST"), by = c("lon", "lat"))
CMC_fill <- right_join(CMC, filter(pixel_match, product == "CMC"), by = c("lon", "lat"))

# Clean up some RAM space
rm(OISST, CMC); gc()

# All the code below can be viewed in upwell_IDX.Rmd
# Next to run upwelling index
# UI Created using SAWS wind data (Find in upwell_IDX_Rmd)
load("Data/UI_angle.RData")

upwelling <- UI_angle %>% 
  dplyr::rename(temp = ui.saws) %>%
  group_by(site) %>%
  # mutate(min_t = min(t), 
  #        max_t = max(t)) %>% 
  nest() %>% # apply the following functions to all of the variables in te dataset
  mutate(clim = purrr::map(data, ts2clm, climatologyPeriod = c("1997-01-01", "2015-12-31")), # creating a column of climatologies. Column will be named clim
         # NB: A threshold of 3 appeared to be far to strict
         # purr::map - apllies a function to each element of a vector
         exceed = purrr::map(clim, exceedance, minDuration = 1, threshold = 1)) %>%  #Upwelling cannot be descrbed as an event. Upwelling can last for a few hours. Given that we have daily data, upwelling events minimum duration here will be 1day
  # Detect consecutive days in exceedance of a given threshold.
  # mutate() %>% 
  select(-data, -clim) %>% 
  unnest() %>%
  filter(row_number() %% 2 == 1) %>%
  unnest() %>% # creates a column for each variables
  dplyr::rename(ui.saws = temp) %>% # rename upwelling index vale to temp so that it could work with the function
  select(site, t, ui.saws, exceedance)

# Now applying the Upwelling func
# NB: This only pulls out the event results and not the climatology results
# This is done to keep the output tidy because group_modify() may only create data.frame type outputs
# To create a list output one would use group_map(), 
# but this then loses the labels of which sites etc. the results belong to
detect_event_custom <- function(df){
  res <- detect_event(df, threshClim2 = df$exceedance, minDuration = 3, coldSpells = T)$event
  return(res)
}

ts2clm_custom <- function(df){
  # The climatology base period used here is up for debate...
  # The choice of the 25th percentile threshold also needs to be justified and sensitivty tested
  res <- ts2clm(df, pctile = 25, climatologyPeriod = c("1992-01-01", "2016-12-31"))
  return(res)
}

# Calculate the upwelling event metrics
upwelling_detect_event <- function(df){
  upwell_base <- df %>% 
    dplyr::rename(t = date) %>% 
    group_by(site, product, heading, distance, lon, lat) %>% 
    group_modify(~ts2clm_custom(.x)) %>% 
    left_join(upwelling, by = c("site", "t")) %>%
    filter(!is.na(exceedance)) %>%
    group_by(site, product, heading, distance, lon, lat) %>% 
    group_modify(~detect_event_custom(.x))
  }

OISST_upwell_base <- upwelling_detect_event(df = OISST_fill)
save(OISST_upwell_base, file = "Data/OISST_upwell_base.RData")
CMC_upwell_base <- upwelling_detect_event(df = CMC_fill)
save(CMC_upwell_base, file = "Data/CMC_upwell_base.RData")
```
